---
title: 'Chapter 3: A review of R modeling fundamentals'
author: "Daniel L."
date: "9/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 3.1 AN EXAMPLE

To demonstrate these fundamentals, let’s use experimental data from McDonald (2009), by way of Mangiafico (2015), on the relationship between the ambient temperature and the rate of cricket chirps per minute. Data were collected for two species: O. exclamationis and O. niveus.

```{r}
library(tidyverse)

data(crickets, package = "modeldata")
names(crickets)
```

Plot the temperature on the x-axis, the chirp rate on the y-axis. The plot elements will be colored differently for each species:

```{r}
ggplot(crickets, aes(x = temp, y = rate, col = species)) + 
  #Plot points for each data point and color by species
  geom_point() + 
  # Show a simple linear model fit created separately for each species:
  geom_smooth(method = lm, se = FALSE) +
  labs(x = "Temperature (C)", y = "Chirp Rate (per minute)")
```

R has many functions that are useful inside of formulas. For example, $poly(x, 3)$ creates linear, quadratic, and cubic terms for $x$ to the model as main effects. The splines package also has several functions to create nonlinear spline terms in the formula.

Returning to our chirping crickets, let’s use a two-way interaction model. In this book, we use the suffix _fit for R objects that are fitted models.

```{r}
interaction_fit <- lm(rate ~ (temp + species)^2, data = crickets);
interaction_fit
```

Before going into any inferential results for this model, the fit should be assessed using diagnostic plots. We can use the plot() method for lm objects. This method produces a set of four plots for the object, each showing different aspects of the fit. Two plots are shown here:

```{r}
# Place two plots next to one another:
par(mfrow = c(1,2))

# Show residuals vs predicted values:
plot(interaction_fit, which = 1)

# A normal quantile plot on the residuals:
plot(interaction_fit, which = 2)
```

Our next order of business with the crickets is to assess if the inclusion of the interaction term is necessary. The most appropriate approach for this model is to re-compute the model without the interaction term and use the anova() method.

```{r}
# Fit a reduced model:
main_effect_fit <-  lm(rate ~ temp + species, data = crickets);

# Compare the two:
anova(main_effect_fit, interaction_fit)
```
This statistical test generates a p-value of 0.25. This implies that there is a lack of evidence for the alternative hypothesis that the interaction term is needed by the model (we fail to reject the null hypothesis). For this reason, we will conduct further analysis on the model without the interaction.

We can use the summary() method to inspect the coefficients, standard errors, and p-values of each model term:

```{r}
summary(main_effect_fit)
```

The chirp rate for each species increases by 3.6 chirps as the temperature increases by a single degree. This term shows strong statistical significance as evidenced by the p-value. The species term has a value of -10.07. This indicates that, across all temperature values, O. niveus has a chirp rate that is about 10 fewer chirps per minute than $O. exclamationis$. Similar to the temperature term, the species effect is associated with a very small p-value.

The only issue in this analysis is the intercept value. It indicates that at 0 C, there are negative chirps per minute for both species. While this doesn’t make sense, the data only go as low as 17.2 C and interpreting the model at 0 C would be an $extrapolation$. This would be a bad idea. That being said, the model fit is good within the $applicable$ range of the temperature values; the conclusions should be limited to the observed temperature range.

If we needed to estimate the chirp rate at a temperature that was not observed in the experiment, we could use the $predict()$ method. It takes the model object and a data frame of new values for prediction. For example, the model estimates the chirp rate for $O. exclamationis$ for temperatures between 15 C and 20 C can be computed via:


```{r}
new_values <- data.frame(species = "O. exclamationis", temp = 15:20)
predict(main_effect_fit, new_values)
```


```{r}
```


