---
title: 'Chapter 4: The Ames housing data'
author: "Daniel L."
date: "10/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Our goal for these data is to predict the sale price of a house based on its other characteristics.

```{r}
library(tidymodels)
data(ames, package = "modeldata")
```

```{r}
dim(ames)
```

## 4.1 EXPLORING IMPORTANT FEATURES

It makes sense to start with the outcome we want want to predict: the last sale price of the house (in USD):

```{r}
#`tidymodels_prefer()` uses the `conflicted` package to handle common conflicts with tidymodels and other packages.
tidymodels_conflicts()

ggplot(data = ames, aes(x = Sale_Price)) + 
  geom_histogram(bins = 50)
```

```{r}
summary(ames$Sale_Price)
```
 
The advantages of doing this are that no houses would be predicted with negative sale prices and that errors in predicting expensive houses will not have an undue influence on the model. Also, from a statistical perspective, a logarithmic transform may also stabilize the variance in a way that makes inference more legitimate. Let’s visualize the transformed data:

```{r}
ggplot(ames, aes(x = Sale_Price)) + 
  geom_histogram(bins = 50) +
  scale_x_log10()
```

While not perfect, this will probably result in better models than using the untransformed data. The downside to transforming the outcome is mostly related to interpretation.

If the sale price is on the log scale, these differences (i.e. the residuals) are also in log units. For this reason, it can be difficult to understand the quality of a model whose RMSE is 0.15 log units.

Despite these drawbacks, the models used in this book utilize the log transformation for this outcome. From $\textbf{this point on}$, the outcome column is pre-logged in the $ames$ data frame:

```{r}
ames <- ames %>% mutate(Sale_Price = log10(Sale_Price))
```

```{r}
summary(ames$Sale_Price)
```

Another important aspect of these data for our modeling are their geographic locations. This spatial information is contained in the data in two ways: a qualitative Neighborhood label as well as quantitative longitude and latitude data. To visualize the spatial information, let’s use both together to plot the data on a map and color by neighborhood.


This is the code to use for the following chapters.

```{r}
library(tidymodels)
data(ames)
ames <- ames %>% mutate(Sale_Price = log10(Sale_Price))
```

```{r}
#
```


