---
title: "Chapter_2"
author: "Daniel L."
date: "9/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
arrange(.data = mtcars, gear, mpg)
```

### 2.1.2 REUSE EXISTING DATA STRUCTURES

As an example, the rsample package can be used to create resamples of a data set, such as cross-validation or the bootstrap (described in Chapter 10). 

```{r}
boot_samp <- rsample::bootstraps(mtcars, times = 3); 
boot_samp
class(boot_samp)
```

One important feature in the tibble produced by rsample is that the $splits$ column is a list. In this instance, each element of the list has the same type of object: an $rsplit$ object that contains the information about which rows of $mtcars$ belong in the bootstrap sample. List columns can be very useful in data analysis and, as will be seen throughout this book, are important to tidymodels.

### 2.1.3 DESIGN FOR THE PIPE AND FUNCTIONAL PROGRAMMING

The magrittr pipe operator (%>%) is a tool for chaining together a sequence of R functions.

```{r}
small_mtcars <- arrange(mtcars, gear)
small_mtcars <- slice(small_mtcars, 1:10)

# or more compactly:
small_mtcars <- slice(arrange(mtcars, gear), 1:10)
```

Using the pipe operator:

```{r}
small_mtcars <- 
  mtcars %>% 
  arrange(gear) %>% 
  slice(1:10)
```

```{r}
library(ggplot2)
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() + 
  geom_smooth(method = lm)
```

