---
title: "Chapter_2"
author: "Daniel L."
date: "9/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
arrange(.data = mtcars, gear, mpg)
```

### 2.1.2 REUSE EXISTING DATA STRUCTURES

As an example, the rsample package can be used to create resamples of a data set, such as cross-validation or the bootstrap (described in Chapter 10). 

```{r}
boot_samp <- rsample::bootstraps(mtcars, times = 3); 
boot_samp
class(boot_samp)
```

One important feature in the tibble produced by rsample is that the $splits$ column is a list. In this instance, each element of the list has the same type of object: an $rsplit$ object that contains the information about which rows of $mtcars$ belong in the bootstrap sample. List columns can be very useful in data analysis and, as will be seen throughout this book, are important to tidymodels.

### 2.1.3 DESIGN FOR THE PIPE AND FUNCTIONAL PROGRAMMING

The magrittr pipe operator (%>%) is a tool for chaining together a sequence of R functions.

```{r}
small_mtcars <- arrange(mtcars, gear)
small_mtcars <- slice(small_mtcars, 1:10)

# or more compactly:
small_mtcars <- slice(arrange(mtcars, gear), 1:10)
```

Using the pipe operator:

```{r}
small_mtcars <- 
  mtcars %>% 
  arrange(gear) %>% 
  slice(1:10)
```

```{r}
library(ggplot2)
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() + 
  geom_smooth(method = lm)
```
Suppose you are interested in the logarithm of the ratio of the fuel efficiency to the car weight. To those new to R and/or coming from other programming languages, a loop might seem like a good option:

```{r}
n <- nrow(mtcars)
ratios <- rep(NA_real_, n)
for (car in 1:n) {
  ratios[car] <- log(mtcars$mpg[car]/mtcars$wt[car])
}
```

```{r}
head(ratios)
```

Those with more experience in R may know that there is a much simpler and faster vectorized version that can be computed by:
```{r}
ratios <- log(mtcars$mpg/mtcars$wt)
```

When we design for functional programming, it is important that the output only depends on the inputs and that the function has no side effects. Violations of these ideas in the following function are shown with comments:

```{r}
compute_log_ratio <- function(mpg, wt) {
  log_base <- getOption("log_base", default = exp(1)) # gets external data
  results <- log(mpg/wt, base = log_base)
  print(mean(results))                                # prints to the console
  done <<- TRUE                                       # sets external data
  results
}
```

A better version would be:

```{r}
compute_log_ratio <- function(mpg, wt, log_base = exp(1)) {
  log(mpg/wt, base = log_base)
}
```

The most basic function, map(), always returns a list and uses the basic syntax of map(vector, function). For example, to take the square-root of our data, we could:

```{r}
map(head(mtcars$mpg,3), sqrt)
```

```{r}
map_dbl(head(mtcars$mpg,3), sqrt)
```

There are also mapping functions that operate across multiple vectors:
```{r}
ratios <- map2_dbl(mtcars$mpg, mtcars$wt, compute_log_ratio)
head(ratios)
```


The map() functions also allow for temporary, anonymous functions defined using the tilde character. The argument values are .x and .y for map2():
```{r}
map2_dbl(mtcars$mpg, mtcars$wt, ~ log(.x/.y)) %>% 
  head()
```

## 2.2 Examples of Tidyverse Syntax

Tibbles have slightly different rules than basic data frames in R. For example, tibbles naturally work with column names that are not syntactically valid variable names:

```{r}
tbbl <- tibble("variable 1" = 1:2, two = 3:4)
tbbl  
```

To demonstrate some syntax, let’s use tidyverse functions to read in data that could be used in modeling. The data set comes from the city of Chicago’s data portal and contains daily ridership data for the city’s elevated train stations. 

```{r, error = FALSE}
library(lubridate)
```

```{r}
url <- "http://bit.ly/raw-train-data-csv"

all_stations <- 
  # Step 1: Read in the data.
  read_csv(url) %>% 
  # Step 2: filter columns and rename stationname
  dplyr::select(station = stationname, date, rides) %>%
  # Step 3: Convert the character date field to a date encoding.
  # Also, put the data in units of 1K rides
  mutate(date = mdy(date), rides = rides / 1000) %>% 
  # Step 4: Summarize the multiple records using the maximum.
  group_by(date, station) %>% 
  summarize(rides = max(rides), .groups = "drop")
```

